---
title: "Creating a nonogram puzzle from an image"
author: "mikefc"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
suppressPackageStartupMessages({
  library(dplyr)
  library(raster)
  
  library(nonogram)
})
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7, fig.height=6
)
```


## Load a small black and white image


```{r}
im <- png::readPNG("figures/Rlogo-bw.png")
plot(as.raster(im), interpolate=FALSE)
```


## From the image, create an integer matrix with just zeros and ones


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Manipulating the image data to select just the 'Green' plane from the
# RGB image, and then thresholding it and inverting it.
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
mat <- 1L - (im[,,2] > 0)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# nonogram has a method for plotting a matrix in a standard way
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
nonogram::create_matrix_plot(mat)
```


## Create the puzzle from the matrix

Puzzles in `nonogram` are represented in 2 ways:

* `puzzles` - list objects with vectors of clues for each row and each column
* `puzzle_strings` - compact character representations of the puzzles
    * a clue for a row is represented by a comma-separated list of integers
    * multiple clues are separated by a colon `:`
    * the row clues come first and then the column clues. 
    * the row and column clues are separated by a `-`


```{r}
puzzle        <- create_puzzle_from_matrix(mat)
puzzle_string <- convert_puzzle_to_puzzle_string(puzzle)
```

```{r echo=FALSE, comment=NA}
cat("> puzzle")
deparse(puzzle) %>% paste(collapse="\n") %>% cat()

cat('> puzzle_string')
strwrap(puzzle_string, width = 60) %>% cat
```


## Create a plot of the matrix and clues 

```{r}
create_puzzle_plot(puzzle, mat, show_clues=TRUE)
```



# Create a blank plot of just the clues

```{r}
create_puzzle_plot(puzzle, show_clues=TRUE)
```


