---
title: "Creating a nonogram puzzle"
author: "mikefc"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
suppressPackageStartupMessages({
  library(dplyr)
  library(magick)
  
  library(nonogram)
})
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7, fig.height=6
)
```

# Create your own puzzles

To make, print and solve your own nonograms, you just need to create a *puzzle string*.

The `puzzle string` format used to define puzzles is quite simple:

* the numbers for each clue are separated by a comma
* each clue is separated by a colon
* the clues for the rows come first, then a dash, then the clues for the columns
* row clues are read from left to right
* column clues are read from top to bottom


```{r}
puzzle_string <- "3:1:1,1-3:1:1,1"
solution      <- solve_puzzle(puzzle_string)
create_puzzle_plot(puzzle_string, solution)
```



# Create your own puzzles from an image

## Load a small black and white image


```{r}
library(magick)
im <- magick::image_read(system.file("img", "Rlogo.jpg", package="jpeg"))


im <- im %>% 
  image_quantize(max=2, colorspace = 'gray', dither=TRUE) %>%
  image_scale(geometry = geometry_size_pixels(width=25, height=20, preserve_aspect=FALSE)) 
  
im
```



## From the image, create an integer matrix with just zeros and ones


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Manipulating the image data to to matrix and then threshold, invert + tranpose
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
mat <- t(1L - 1L * (im[[1]][1,,] > 180))

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# nonogram has a method for plotting a matrix in a standard way
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
nonogram::create_matrix_plot(mat)
```


## Create the puzzle from the matrix

Puzzles in `nonogram` are represented in 2 ways:

* `puzzles` - list objects with vectors of clues for each row and each column
* `puzzle_strings` - compact character representations of the puzzles
    * a clue for a row is represented by a comma-separated list of integers
    * multiple clues are separated by a colon `:`
    * the row clues come first and then the column clues. 
    * the row and column clues are separated by a `-`


```{r}
puzzle        <- create_puzzle_from_matrix(mat)
puzzle_string <- convert_puzzle_to_puzzle_string(puzzle)
```

```{r echo=FALSE, comment=NA}
cat("> puzzle")
deparse(puzzle) %>% paste(collapse="\n") %>% cat()

cat('> puzzle_string')
strwrap(puzzle_string, width = 60) %>% cat
```


## Create a plot of the matrix and clues 

```{r}
create_puzzle_plot(puzzle, mat, show_clues=TRUE)
```



## Create a blank plot of just the clues

```{r}
create_puzzle_plot(puzzle, show_clues=TRUE)
```


