---
title: "Nonogram solution technique"
author: "mikefc"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
suppressPackageStartupMessages({
  library(dplyr)
  library(nonogram)
})

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



Nonograms Solution Technique
------------------------------------------------------------------------------

This vignette outlines the solutino technique used in this package.


Nomenclature
------------------------------------------------------------------------------


### `clue`

A clue is sequence of integers which are the run-length encoding of the filled-in 
squares in a row or column. 

E.g. **`1,2`** is a *clue* which means that there is 1 filled-in square in this row
  (or column) followed by some blank squares, and then two filled-in squares in a row
  
### `pattern`

A *pattern* is a sequence of filled-in and blank squares which make up a row or 
column.  

E.g. the clue `1,2`  (on a board 8 squares wide) is satisfied by each
of the following patterns, represented as a matrix and shown graphically

```{r echo=FALSE}
pattern_set <- nonogram::create_pattern_set(c(1, 2), 8)
pattern_set <- pattern_set[c(1, 5, 11), ]

pattern_set

nonogram::create_matrix_plot(pattern_set, row_spacing = 0.3)
```


### `pattern set`

A *pattern set* is the complete set of all possible patterns which satisfy a clue.

E.g. the pattern set for the clue `2,3` on a board 8 squares wide is shown below
as a matrix as well as graphically.

```{r echo = FALSE}
pattern_set <- nonogram::create_pattern_set(c(2, 3), 8)
pattern_set
nonogram::create_matrix_plot(pattern_set, row_spacing = 0.3)
```


### `pattern sets`

For each clue along the rows, and each clue along the columns, there is a *pattern set*.

The *row pattern sets* is the collection of the *pattern set*s for every row.

The *column pattern sets* is the collection of the *pattern set*s for every column.


### `puzzle`

A puzzle is collection of clues, represented in the `nonogram` package as a list of
vectors.

E.g. the puzzle for the **duck** image shown at the top is:

```{r eval=FALSE}
puzzle <- list(
  row_clues = list(3L, 2:1, 3:2, c(2L, 2L), 6L, c(1L, 5L), 6L, 1L, 2L), 
  col_clues = list(1:2, c(3L, 1L), c(1L, 5L), c(7L, 1L), 5L, 3L, 4L, 3L)
)
```



### `puzzle string`

For storage and transfer, it's sometimes easier to deal with the puzzle just
as a string.

A *puzzle string* is a compact string representation of the puzzle.

It consists of:

* the numbers for each clue separated by a comma
* each clue separated by a colon
* the row and column clues (in that order) separated by a dash


As an example, the following is the puzzle string for the **duck** nonogram



```{r eval=FALSE}
"3:2,1:3,2:2,2:6:1,5:6:1:2-1,2:3,1:1,5:7,1:5:3:4:3"
```




Solution technique
------------------------------------------------------------------------------

1. For each clue generate its pattern set
2. Within each pattern set, find if any locations are a constant colour, and 
   use this fact to elimintate patterns in the other direction
3. On the remaining (hopefully small) number of choices within each pattern set
   for each clue, do a recursive search with back-tracking to find a solution
   
   
Solution step 1: Generate all possible patterns for a clue
------------------------------------------------------------------------------

This was the most difficult part of this solution technique, as the number of 
possible combinations, and the resulting `pattern set`, can be **huge**.

Most grids in print max out at about 35 squares along each side.  There are plenty
of larger ones, but I optimised the solver only up to this size.  

For a small puzzle, like the **duck** a the clue in the first column (`1,2`) has
21 possible patterns I've plotted them below

```{r fig.height=3, echo=FALSE, fig.width=6}
pattern_set <- create_pattern_set(c(1, 2), 9)

create_matrix_plot(t(pattern_set), col_spacing = 0.45)
```


For the larger `gchq` puzzle below, the clue `1,1,2,1,1,2` to fit in the
puzzle width of 25 squares has 18,564 patterns, some of which are shown below.

```{r fig.height=3, echo=FALSE, fig.width=6}
pattern_set <- create_pattern_set(c(1,1,2,1,1,2), 25)
pattern_set <- pattern_set[c(1, 10, 100, 200, 300, 1000, 10000, 15000),]

create_matrix_plot(pattern_set, row_spacing = 0.4)
```


For a clue of `1,1,2,1,1,2,1,1,1,1` in a puzzle that is 35 squares wide, there 
are **1,961,256** patterns.


In the end, I've settled upon the following solution:

* recursively build the solution
* with back-tracking whenever a valid pattern is impossible from a particular location
* store all values as raw matrices (to reduce the memory pressure)
* have a cache of patterns for smaller clues and puzzle widths rather than recalcualting every time.


If I ever revisit this problem, I have an idea for a solution using the 
[memoise package](https://cran.r-project.org/package=memoise).  But the problem
with cacheing *everything* is that some of the pattern sets are *HUGE* - more than 
a gigabyte - and I don't want to cache all those.

If I ever wanted to solve bigger nonograms,
I'd need to use another solution technique altogether - the up-front generation of all
possible clues just takes up too much time and memory.


Solution step 2: Filter the pattern sets
------------------------------------------------------------------------------

For very *very* small nonograms, it might be possible to try every combination 
of patterns in a row and exhaustively search for a solution.

In the **duck** problem the total number of combinations for all patterns for all
rows is 8x10^6 (8 million).  If you could try 10,000 patterns/sec, you could 
try them all in about 15 minutes.

For some of the larger nonograms there are more than 1x10^100 (not a typo) combinations to try if you
wanted to brute force a solution!

To initally filter the set of all patterns into a more tractable solution, I 
look for locations within each pattern set which can only be one colour and use that to 
filter the orthogonal data set.


E.g. The pattern set for the third row in the **duck** puzzle with clue `3,2` is shown below:

```{r echo=FALSE, fig.height=3}
puzzle_string_examples[['duck']] %>% convert_puzzle_string_to_puzzle() -> puzzle
create_pattern_set(puzzle$row_clues[[3]], 8L) %>%
  create_matrix_plot(row_spacing = 0.3)
```

The third column is always filled-in! Which means that I now look at the pattern
sets for the third column and keep only ones with that location filled in.

The clue for the third column is `1,5` which looks like:

```{r echo=FALSE, fig.height=3}
pattern_set <- create_pattern_set(puzzle$col_clues[[3]], 9L) 
create_matrix_plot(t(pattern_set), col_spacing = 0.3)
```

Only two patterns have the required filled-in square in the 3rd row, the others
can be eliminated from consideration.

This process is repeated until there are no more occurrences.

Sometimes by doing this filtering there is only one possible pattern left for
each row and column, in which case we have found a solution!


Solution step 3: Recursively build the solution by rows with back-tracking 
------------------------------------------------------------------------------

If there are still multiple patterns to try for each row we now try all of them
in turn by choosing them one row at a time.

But as each row is selected, a cross-check is done to see that at least one 
pattern in each column can fit that sequence of rows.

If it fails then we stop that search path, select a different
row from the pattern set and continue.  

